---
title: "stl_lead_inequity.qmd"
format: html
editor: visual
---

### Library chunk

```{r}
library(janitor)
library(tidyverse)
library(skimr)
library(here)
```

### Read in Data and clean up

```{r}
stl_lead <- read_csv(here::here("data", "stl_blood_lead_commas.csv")) %>% 
  clean_names()
```

### Basic Exploration

```{r}
head(stl_lead)
tail(stl_lead)
skim(stl_lead)
```

### Create new column that is the proportion of white entries 

```{r}
stl_lead_prop <- stl_lead %>%
  mutate(prop_white = white / total_pop * 100)

#---this is going to 1. create and store a new data frame, 2. that new data frame is the old one piped into the mutate function which we tell to add a new column which we define accordingly
```

### Create graph of those in each census tract with elevated lead versus those identifying as white 

is there a way to change the dimensions of this plot within the html render? seems to be ignoring the height and width arguments

```{r}

stl_lead_plot <- ggplot(data = stl_lead_prop, aes(x = pct_elevated, y = prop_white), height = 7, width = 1) +
  geom_point(aes(color = pct_elevated))
stl_lead_plot

```

we are noticing what appears to be (without actually determining a level of correlation) a negative correlation between those identifying as white and those with elevated lead in their blood. This means that as the elevated lead levels increase, the population proportion white decreases

### Create histogram of pct elevated 

```{r}
elevated_hist <- ggplot(data = stl_lead_prop, aes(x = pct_elevated)) + geom_histogram(binwidth = 1)
elevated_hist
```
